Staged Abstract Interpreters

### 2019-04-05 paper/60_cases.tex
evaluation

We first revisit a similar technique called \textit{abstract compilation} (AC).
AC is introduced by \citet{Boucher:1996:ACN:647473.727587} as a new
implementation technique for abstract interpretation based static analysis.
Similar to the present paper, the idea is inspired by partial evaluation:
the program is known statically and the interpretive overhead can be
eliminated. In AC, the compiled analysis can be represented by either text or
closures (higher-order functions). The minute difference is that the closures
can be executed immediately, while the textual programs need to be compiled and
loaded.

@@

On the other side, our approach does exactly the same thing through staging: the
AST are static, and stores are dynamic, therefore we annotate the type @Store@
with @Rep@ (shown in Figure \ref{compare_ac_sai} (right)). In fact, the only
changes of our appraoch using LMS framework are the types of @Store@ to
@Rep[Store]@, since they will be known at the next stage. The generated code
will just looks-up and update the store. The functions at current stage are
eliminated in the residual program, therefore we do not generate higher-order
functions, which provides additional performance improvement. Nevertheless,
the support from LMS framework is necessary, such as staging support for @Map@,
we consider these efforts are reusable and complete orthogonal to the
implementation of analysis.

>>>>

We first revisit a similar technique called \textit{abstract compilation} (AC).
AC was introduced by \citet{Boucher:1996:ACN:647473.727587} as an
implementation technique for abstract interpretation based static analysis.
Similar to the present paper, the idea is inspired by partial evaluation:
the program is known statically and the interpretive overhead can be
eliminated. In AC, the compiled analysis can be represented by either text or
closures (higher-order functions). The minute difference is that the closures
can be executed immediately, while the textual programs need to be compiled and
loaded.

@@

On the other side, our approach does exactly the same thing through staging: the
AST are static, and stores are dynamic, therefore we annotate the type @Store@
with @Rep@ (shown in Figure \ref{compare_ac_sai} (right)). In fact, the only
changes of our appraoch using LMS framework are the types of @Store@ to
@Rep[Store]@, since they will be known at the next stage; other parts do not need to change. 
The generated code
will just looks-up and update the store. The functions at current stage are
eliminated in the residual program, therefore we do not generate higher-order
functions, which provides additional performance improvement. Nevertheless,
the support from LMS framework is necessary, such as staging support for @Map@,
we consider these efforts are reusable and complete orthogonal to the
implementation of analysis.

### 2019-04-05 paper/70_eval.tex
evaluation

\paragraph{Setting}

We use Scala 2.12.4 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
(kernel 4.4.0) machine. All of our evaluation were performed on a machine with 4 Intel
Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96
cores (192 threads) in total, the abstract interpreters are single-threaded.
To minimize the effect of warming-up from the JVM JIT compiler, all the
experiments are executed for 20 times and we report the statistical median value
of the running times (in seconds).

\paragraph{Benchmarks} \todo{Rewrite}
The benchmark programs we used are collected from previous papers
\cite{Johnson:2013:OAA:2500365.2500604, ashley:practical,
DBLP:journals/corr/abs-1102-3676} and exisiting artifacts
\footnote{https://github.com/ilyasergey/reachability} for control-flow analysis.
We used the following benchmark programs:

\begin{itemize}
  %\item \textbf{fib}: Calculates the $n$-th fibonacci number.
  \item \textbf{rsa}: The RSA public key encryption algorithm.
  \item \textbf{kcfa3}: A difficult benchmark for $k$-CFA.
  \item \textbf{church}: Church numerals with additions and mutiplications.
  \item \textbf{fermat}: Fermat and Solovay-Strassen primality testing.
  \item \textbf{kcfa-worst-case-}\textit{n}: 
    Benchmark programs that are supposed to be tough cases for $k$-CFA; 
    the number $n$ indicates the depth of nesting lambda terms.
\end{itemize}

@@

\begin{figure}[h]
\tiny
\begin{tabular}{@{}lllll@{}}
\toprule
    Program          &\#AST & unstaged   & staged     & $\Delta$ \\ \midrule
    @rsa@            &      &            &            & \%          \\
    @kcfa3@          &      &            &            & \%          \\
    @church@         &      &            &            & \%          \\
    @fermat@         &      &            &            & \%          \\
    @kcfa-worst-16@  &  198 & 0.352      & 0.028      & +1157.14\%        \\
    @kcfa-worst-32@  &  390 & 4.234      & 0.289      & +1365.05\%          \\
    @kcfa-worst-64@  &  774 & 61.475     & 4.425      & +1289.27\%         \\
    \bottomrule
\end{tabular}
\caption{Evaluation result for 0CFA without store-widening.} \label{evaluation_result}
\end{figure}

>>>>

\paragraph{Setting}
We use Scala 2.12.7 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
(kernel 4.4.0) machine. All of our evaluation were performed on a machine with 4 Intel
Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96
cores (192 threads) in total, the abstract interpreters are single-threaded.
To minimize the effect of warming-up from the JVM JIT compiler, all the
experiments are executed for 20 times and we report the statistical median value
of the running times (in seconds).

\paragraph{Benchmarks}
The benchmark programs we used in the experiment are collected from several previous papers
\cite{Johnson:2013:OAA:2500365.2500604, ashley:practical,
DBLP:journals/corr/abs-1102-3676} and exisiting artifacts
\footnote{https://github.com/ilyasergey/reachability} for control-flow analysis.
Some of the benchmarks are small programs designed for experiments, such as the @kcfa-worst-@$n$ series, 
which are intended to challenge $k$-CFA; while some other benchmarks are from real-world applications, 
for examples, the RSA public key encryption algorithm @rsa@ and matrix computation @matrix@.
In the Table \ref{?}, we report the number of AST node after parsing and desugaring 
(excluding comments) as a proper measurement of program size.

\iffalse
We used the following benchmark programs:
\begin{itemize}
  %\item \textbf{fib}: Calculates the $n$-th fibonacci number.
  \item \textbf{rsa}: The RSA public key encryption algorithm.
  \item \textbf{kcfa3}: A difficult benchmark for $k$-CFA.
  \item \textbf{church}: Church numerals with additions and mutiplications.
  \item \textbf{fermat}: Fermat and Solovay-Strassen primality testing.
  \item \textbf{kcfa-worst-case-}\textit{n}: 
    Benchmark programs that are supposed to be tough cases for $k$-CFA; 
    the number $n$ indicates the depth of nesting lambda terms.
\end{itemize}
\fi

@@

\begin{figure}[h]
\tiny
\begin{tabular}{@{}lllll@{}}
\toprule
    program          &\#AST & unstaged   & staged     & $\frac{\text{unstaged}}{\text{staged}}$ \\ \midrule
    @church@         & 120  & 71.811     & 23.990     & 299.34\%          \\
    @rsa@            & 451  & 0.308      & 0.064      & 459.15\%          \\
    @lattice@        & 609  & 31.874     & 8.816      & 358.69\%          \\
    @fermat@         & 310  & 0.043      & 0.006      & 716.67\%          \\
    @mbrotZ@         & 331  & 1.112      & 0.214      & 519.63\%          \\
    @kcfa-worst-16@  & 182  & 0.083      & 0.013      & 638.46\%          \\
    @kcfa-worst-32@  & 358  & 0.452      & 0.044      & 1027.27\%         \\
    @kcfa-worst-64@  & 774  & 3.329      & 0.236      & 1410.59\%         \\
    \bottomrule
\end{tabular}
\caption{Evaluation result for 0CFA without store-widening.} \label{evaluation_result}
\end{figure}
